Bayesian exponential random graph modeling of KWL Burial network

### 2020-10-15 analysis/paper/paper.Rmd
edit method section and work on egodata

<!--burial study- network analysis-->
Burials contains valuable information about individual mortuary practices and inherited structure of a society as a whole. With the development of network science and graph theory, it is promising to study relation among burials and explore its social implication [@Sosna2013]. Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow archaeologists to infer social structures. This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015]. For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom. Burials with grave goods in common often indicates underlying social relation connected through trade, exchange, and gifting [@Coward2013, pp. 252]. Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization. A common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies recently [@Mills2017]. Networks are represented by graphs that consists of a set of actors, such as people, groups, or organizations, and a set of links connecting them. Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes. The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016]. In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relations in societies [@Marin2011]. With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affects the structure [@Freeman2004; @Salvini2010]. The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]

@@

<!--current approaches to network analysis-->
Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures by directly modeling the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that an observed network is randomly generated depending on the presence or absence of other edges. ERGMs differs from traditional network approaches that edges are only formed according to their corresponding predictors separately without taken into account the interactions among edges [@Morris2008]. In statistical models, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011]. For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby. Similarly, for network dependencies, reciprocity represents the likelihood between two actor's to be tied because of shared properties, and transitivity means two actors tend to be linked via a relationship with a third actor in common. We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007]. Network specification under assumptions and contexts is useful for exploring archaeological network. For instance, @Amati2019 demonstrated how to reconstruct of three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms using ERGMs. Three network models test different hypotheses for the relationship between sites based on proximity, inter-cultural, and exchange network to explain the presence of hub sites they observed. They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one model exclusively. However, they also pointed out some limitations of ERGMs, such as static outcome and sensitivity to missing data. In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits the widely use of ERGMs. Here, we used ERGMs in a Bayesian framework that allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011]. MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].

@@

## data and hypoethesis

We analyzed archaeological burials from an Iron Age site, Kiwulan, in northeastern Taiwan based on data collected from published excavation report, and original fieldwork notes to generate networks. We do not include the burials in Chinese phase in our network analysis because of the small sample size. A network is a structure can be represented as a graph consisting of a set of nodes (actors) and edges (ties) connecting them [@Brughmans2018; @Robins2007]. In our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common. Due to gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels based on the distribution of their counts to indicate quantity, high, medium, and low. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For prestige goods not commonly found across burials, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess the good in common. Node attributes here include osteological data and mortuary practices, such as age, gender, burial ritual, and wealth level. Burial ritual is identified by locally made ceramics on the top of grave that suggests the presence of funeral feasting. Wealth level is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors.

>>>>

<!--burial study- network analysis-->
Burials contains valuable information about individual mortuary practices and inherited structure of a society as a whole. With the development of network science and graph theory, it is promising to study relation among burials and explore its social implication [@Sosna2013]. Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow archaeologists to infer social structures. This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015]. For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom. Burials with grave goods in common often indicates underlying social relation connected through trade, exchange, and gifting [@Coward2013, pp. 252]. Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization. A common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies recently [@Mills2017]. Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007]. Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes. The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016]. In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relations in societies [@Marin2011]. With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affects the structure [@Freeman2004; @Salvini2010]. The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]

@@

<!--current approaches to network analysis-->
Exponential Random Graph Models (ERGMs) are an important family of statistical models for networks that allows directly modeling for the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that an observed network is randomly generated depending on the presence or absence of other edges. ERGMs differs from traditional network approaches that edges are only formed according to their corresponding predictors separately without taken into account the interactions among edges [@Morris2008]. In statistical models, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011]. For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby. Similarly, for network dependencies, reciprocity represents the likelihood between two actor's to be tied because of shared properties, and transitivity means two actors tend to be linked via a relationship with a third actor in common. We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007]. Network specification under assumptions and contexts is useful for exploring archaeological network. For instance, @Amati2019 demonstrated how to reconstruct of three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms using ERGMs. Three network models test different hypotheses for the relationship between sites based on proximity, inter-cultural, and exchange network to explain the presence of hub sites they observed. They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one model exclusively. However, they also pointed out some limitations of ERGMs, such as static outcome and sensitivity to missing data. In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits the widely use of ERGMs. Here, we used ERGMs in a Bayesian framework that allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011]. MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].

@@

## data and hypothesis

We analyzed burial data collected from published excavation reports, and original fieldwork notes from Kiwulan. We do not include the burials of the Chinese phase in our network analysis due to small sample size. In our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common. Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess each type of goods in common. Node attributes here include osteological data and mortuary practices, such as age, gender, presence of burial rituals, and overall burial goods quantity index. Burial ritual is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records. Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.
